\section{Introduction}

The goal of this chapter is to organize the issues people have with
{\bibtex}: answering what the issues in {\bibtex} are
\chapref{sec:intro_what_issues}, and how to approach the {\bibtex}
issues \chapref{sec:intro_what_to_do}.


\section{What are the issues in {\bibtex}}
\label{sec:intro_what_issues}

{\bibtex} has changed the landscape for scientific writing.  However,
not without any issues.  As seen in
Chapter~\ref{ch:problem-description}, the challenges range widely and
trying to group similar looking issues we have:

The misspellings in general, misspellings in names, initials in author
names, erroneous online lookups, use of abbreviations for journal
names, and {\bibtex} strings that end up being text will be considered
as lexical concerns.  Conforming to the specification and de-facto
standards is considered as a correctness concern.  All of these issues
will be considered combined as \newdef{correctness and lexical
  concerns}.

Duplicated values, forum names that change, inconsistent use of tags
and inconsistent entry keys will be considered as \newdef{consistency
  concerns}.

The Utopian goal is to provide a structural solution for all these
issues so that if any further issues exist, they would be purely
conjunctural.


\section{What can be done about the {\bibtex} issues}
\label{sec:intro_what_to_do}

As previously stated, the structural approach to the issues is
desired.  This choice means to ensure that there are tools that are
able to handle the issues - preferably to the level where all issues
are solved.  As touched upon shortly when inspecting the problems, it
is not likely that all issues can be solved perfectly.  For a
structural solution, there are two approaches:


\subsection{Updating or replacing {\bibtex}}

One way of handling the issues structurally would be to change or
replace {\bibtex}, so it handles all lexical and consistency concerns.
This way would include changing the {\bibtex} specification to account
for relevant de facto standards, enforcing conformity, handling
abbreviations and controlling all data.  The updated version of
{\bibtex} could then either correct the issues when running into them
or fail building the \file{bbl} with appropriate error messages for
issues that the user needs to take care of.

This approach would probably be perceived as invasive since it would
cause existing {\bibtex} files not to work, and it would impose
requirements one may not desire.  The perception would of course
depend on perspective because the user who wants structure and control
might find it good that it is enforced.

As will be inspected in Chapter~\ref{ch:related} there are actually a
few attempts at both changing and replacing {\bibtex}.


\subsection{Augmenting {\bibtex}}

Instead of changing or replacing {\bibtex}, an augmenting tool is
another option.  Such a tool, together with {\bibtex}, would provide
or suggest improvements, instead of changing specifications.  An
augmenting tool will be a supplement to current use of {\bibtex} and
be optional, rather than imposed on the users.

An augmenting tool also has the advantage of separating concerns.  The
concern of {\bibtex} is to cite references and print a bibliography
according to a bibliography style.  Where as the concern of an
augmenting tool is to ensure that ones has a nice \file{bib}.

Getting people to use such a solution ironically cause a new issue, to
which there is no structural solution.

\section{How do we approach the {\bibtex} issues}
\subsection{Introduction}

The goal of this section is to introduce our choice of solutions for
the issues.


\subsection{Lexical and correctness concerns vs. consistency concerns}
\label{sec:approach_lexical_consistency}

The relation between the lexical and correctness concerns and the
consistency concerns reveals a dependency in the analysis.  Going
through the consistency concerns observing their relation gives:

\begin{itemize}
\item For inconsistent use of entry tags, it is needed to have a way
  to determine if entries are from the same forum.  Such a
  way depends on consistent naming of the forum and a way to detect
  name changes.

\item For duplicate entries having unique identifiers such as arXiv
  numbers, ISSN or DOI will make the detection trivial.  Otherwise,
  the detection has to be based on the similarity of the information:
  at best the information is identical, otherwise it has to be as
  similar as possible to improve the detection.  Thus solving the
  lexical concerns will be of use.  For repeated content detecting
  when there are multiple instances of $@INBOOK$, $INCOLLECTION$ and
  $INPROCEEDINGS$ that reference the same source is needed.  For
  textual content reused content, such as a journal name, should be
  detected.  For textual values it should compare the values that are
  likely to be repetitive.

\item Inconsistent naming of entry keys can be handled by a naming
  scheme.  Such a naming scheme is usually based on the information in
  the entries.  So having the relevant tags and correct content in
  them will provide a way to ensure consistent entry keys.

\item For name changes of forums, we need to be able to recognize the
  names, which is easier with correct and consistent names.
\end{itemize}

A common property about the consistency concerns is that they are
easier to handle, once the lexical and correctness concerns have been
handled.  This property indicates that a two phase solution may be
desired: first handling the lexical and correctness concerns, then
handling the consistency concerns.


\subsection{\nameref{sec:problems_duplicates}}
\label{sec:approach_duplicates}

Duplicate entries are fairly straight forward if the tags and the
contents are identical.  If the content and tags deviate, a way to
detect ``similarity'' will be needed.  The easiest definition of
similarity is if the title and author is identical, however this
definition might need to take things like revisions and year into
account, as an author may decide to write a new version later or if
one for some reason desires to refer to different revisions.  Further
challenges may arise if there are lexical and correctness issues.  As
per Section~\ref{sec:approach_lexical_consistency}, one should fix
these issues first.

For duplicate values two ways of detecting are needed.  Provided the
correctness of the values, detecting duplicated meta-data is done by
corresponding the $author$/$editor$ and $title$ for $@BOOK$, $@INBOOK$
and $@INCOLLECTION$, for conference proceedings comparing the $title$
of $@PROCEEDINGS$ and the $booktitle$ of $@INPROCEEDINGS$ will
determine if the conference is the same,


\subsection{\nameref{sec:problems_spelling}}

To detect misspellings, a spell checker can be used.  Alternatively,
checking the resources in online databases is an option.  If a spell
checker is used, one should be aware of false positives.
Domain-specific terms might not be present in the dictionary and if
the original source is misspelled, it would be a mistake to correct it
(once published, the name published is the correct name of the
reference!).  Provided a solution for the issues in online lookups,
the correct spelling will be a matter of looking up, but references
may not be in the databases.  As stated in
Section~\ref{sec:approach_look_ups}, there is no good way to ensure
correct lookups.  Therefore a spell checker seems like a good way to
get an indication of possible misspellings, but one would still have
to verify them.  Since entries may be in different languages, a way of
specifying the language should be considered.


\subsection{\nameref{sec:problems_spelling_names}}
\label{sec:approach_spelling_names}

Spell checking names with a normal spell checker is likely to cause
false positives.  A possible way to approach these false positives
would be to widen the online search to databases with scientific
authors, such as DBLP and Google Scholar.  This widening, however,
will not help for authors who are not in these other databases.
Extending this solution to contain more databases such as book authors
will improve the solution, but will still be limited to known author
names.  Another issue is that some authors use different versions of
their name depending on the context.  For example, Iain Menzies Banks
wrote both mainstream fiction and science fiction, writing under the
name Iain Banks for mainstream fiction and using Iain M. Banks for
science fiction~\cite{community2016_banks}.


\subsection{\nameref{sec:problems_initials}}

Finding the initials is a matter of being able to detect single
letters with or without a period after it.  However, if one for some
reason groups initials together, \eg, making George R. R. Martin into
George RR Martin, or J. K. Rowling into JK Rowling, or B. B. King into
BB King, then further detection will be needed.  Replacing the
initials with full names is appropriate whenever possible, but since
the full names may not be known some way of specifying that initials
is the only thing available is needed.  The best approach will
probably be the one described for spell checking names in
Section~\ref{sec:approach_spelling_names}.


\subsection{\nameref{sec:problems_look_ups}}
\label{sec:approach_look_ups}

Online database lookups can be a very useful tool for handling the
lexical and correctness concerns, but getting incorrect data can cause
problems.  The best approach would be to ensure correct lookups, by
using services that are known to be correct.

A situation where relatively reliable lookups is possible, as in the
``EPTCS'' lookup seen in \figref{fig:eptcs_lookup}, can be used to
improve the reliability of the results.  However, there is still no
certain way to know if the database of the service is correct, so it
is still not certain.  Most likely the ID systems, such as arXiv
numbers, DOI and ISSN, will also provide a relatively reliable lookup
mechanism, but it is still not guaranteed.

Another way to approach the bad lookups could be by doing the same
lookup in multiple databases and then have some kind of voting system
that decides on which entry to trust.  This would however require
knowledge information sources each online service use, because their
source of information may be the same and then the same error could
get multiple votes.  The approach can be refined by having increased
trust in databases that are likely to be correct.

The most appropriate strategy is probably selecting the database most
likely to be correct and then have the user select if one agrees with
the result.  Doing the vote system would be overkill in most
situations, and the user would still has to validate the result, since
the voting system will not provide a certain correct result.  Having
the user validate the result will make issues partly conjunctural, if
one just accept any result from the lookup.


\subsection{\nameref{sec:problems_name_changes}}

Handling name changes of forums is supportive to ensuring consistency.
Since name changes cannot be derived automatically, one approach would
be a database of known name changes, which has the disadvantage that
it needs to be maintained.  Adding a configuration to specify name
changes may also be appropriate.


\subsection{\nameref{sec:problems_de_facto}}

As stated in Section~\ref{sec:problems_de_facto} it is desired to be
able to validate if the file conforms to the specification and the
desired de-facto standards.  Validating conformity to the
specification is a simple task, as the specification is just a set of
rules (Section~\ref{sec:about_micro_use}).  A set of de-facto
standards, likewise, is also a simple set of rules.

De-facto standards however, provide challenges, as they are the
standards currently in use.  This means that they both depend on who
the user is and the standards are subject to change.

A tool handling conformity to the specification and de-facto standards
should thus be configurable to account for changes in de-facto
standards.  For practicality, the de-facto standards that are not
likely to change (such as ISSN and DOI) could be accounted for with a
default setting.

Validating the values will improve this solution further.  Having
rules for valid values whenever possible.  Some rules will be very
easy to set up, \eg, ISBN consists only of numbers and dashes.  Other
rules such as not allowing numbers inside a name might be able to
become an issue, \eg., if the queen of Denmark were to write something
that is later cited a correct name for the author is ``Dronning
Magrethe 2.''.  A structural solution will validate as much as
possible, but too rigid validation has the potential of causing new
structural issues.


\subsection{\nameref{sec:problems_abbreviations}}

Ensuring a consistent use of either abbreviations or full names is
desired.  From the point of having the information in a complete
version converting full names, \newdef{de-abbreviating} is desired.
Using a database of standard abbreviations for forums will be useful
to de-abbreviate.  Taking care of a consistent way to switch between
full names and abbreviations is also desired.  Making use of strings
to handle the switching between full names and abbreviations is
probably the best approach since this will keep it clear which forum
is which.  This also allows the user to use string names that are:
full names, official names or their own style of abbreviations, to
their choice.


\subsection{\nameref{sec:problems_strings_as_text}}

A {\bibtex} string can end up being part of a text by mistake: in the
example used in Section~\ref{sec:problems_strings_as_text} where the
month ended up as a text rather than a string a simple check is
possible, because for a month we know what to expect.  Whenever
something in the middle of the text should have been a string, the
text would have to be checked for potential strings.  Automatically
correcting it would introduce a potential source of errors, because a
text being identical to a string name could just be a coincidence, so
it has to be the user's choice.


\subsection{\nameref{sec:problems_inconsistent_tags}}

Detecting inconsistent use of tags requires a way of detecting when
entries are from the same forum.  When such a way is provided, it is
possible to check if the set of tags are the same.  Having some kind
of statistics on the usage may further improve the feedback, since it
will be possible to suggest the shortest path to consistency, whether
by adding or by removing tags.

Since a lot of the forums are continuous, such as a conference being
held each year, the detail level of the information may change over
time.  Also, in some cases, a single item can have additional
information that is not general to the forum, or for some reason not
have information according to the general standard.  Optimally, there
should be a way to account for these cases, either by the user
enforcing conformity, or having options for when a desired deviation.


\subsection{\nameref{sec:problems_inconsistent_keys}}

Provided that the lexical and correctness concerns have been solved,
handling inconsistent entry keys require very little effort.  Having a
rule for how the key names should be formatted is all that is needed.
Like in Section~\ref{sec:approach_duplicates} there is the issue of
similar, but different entries.  Similar entries could result in the
same entry key, so there is the need for a way to disambiguate the key
names.  Since a lot of users already have databases in use, support
for one specifying a naming scheme would be appropriate.


\section{Summary and conclusions}

The issues in {\bibtex} files have been grouped in correctness and
lexical concerns and in consistency concerns.  Updating or replacing
{\bibtex} was compared to augmenting {\bibtex}.  It was observed that
the consistency concerns in general depend on the solution of the
correctness and lexical concerns.

\begin{itemize}
\item Duplicate entries can be found if there is a unique identifier
  or identical entries.  To detect deviating duplicates, a definition
  of ``similarity'' will be needed.

\item Spelling errors in general can be solved by a spell checker and
  the usage of online lookups.  For a spell checker, one must be aware
  of false positives and language.

\item Spelling errors in names will challenge normal spell checkers.
  Using online databases of authors will enable some checking, but the
  solution will be limited to known author names.

\item Initials hiding people's names can be handled by online
  resources, just as misspellings in names.

\item To get online lookups that contain the correct data is
  impossible, the results can be improved by selecting the most
  appropriate database for the lookup and by introducing detection of
  erroneous lookups.

\item Name change of a forum can be handled by a database and by a way
  to specify name changes.

\item Conformity to de-facto standards and the {\bibtex} specification
  should be handled by checking the rules for the standards and being
  able to specify the desired de-facto standards.

\item Journal abbreviations should be moved into strings and be
  consistently de-abbreviated or abbreviated.

\item {\bibtex} strings that end up as part of the text should be
  detected by matching string names that appear in text.

\item Inconsistent tags should be detected based on when entries are
  from the same forum.  One should be able to specify deviations from
  the general set of information for the forum.

\item Inconsistent entry keys should be handled by a rule for the
  desired format for entry key names.
\end{itemize}

Handling the {\bibtex} issues will be done by augmenting {\bibtex}
with a multiple phase tool.  The tool will address the correctness and
lexical concerns first, then the consistency concerns.


%\begin{figure}
%  \centering
%  \begin{verbatim}
%This is BibTeX, Version 0.99d (TeX Live 2015/Debian)
%The top-level auxiliary file: doc.aux
%The style file: plain.bst
%Database file #1: mybib.bib
%Warning--can't use both author and editor fields in book
%Warning--empty publisher in book
%Warning--empty year in book
%Warning--empty journal in Nobody06
%Warning--empty year in Nobody06
%\end{verbatim}
%  \caption{Example of error output from {\bibtex}}
%\label{fig:bibtex_out}
%\end{figure}


%\remark{People who request similar
%http://stackoverflow.com/questions/13630584/the-best-method-for-handling-bibtex-files
%http://stackoverflow.com/questions/32838020/awk-how-to-clean-bibtex-files
%http://www.latex-community.org/forum/viewtopic.php?f=50&t=12043
%http://tex.stackexchange.com/questions/76420/cleaning-up-a-bib-file
%http://tex.stackexchange.com/questions/174509/is-there-a-tool-service-that-can-enrich-a-bibtex-database
%<http://tex.stackexchange.com/questions/128989/how-can-one-validate-a-bib-file
%http://tex.stackexchange.com/questions/173621/how-to-validate-check-a-biblatex-bib-file
%http://stackoverflow.com/questions/13630584/the-best-method-for-handling-bibtex-files
%}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "thesis"
%%% End:
