\documentclass[twoside,11pt,openright,a4paper]{report}

\usepackage{cstitle}
\usepackage[utf8]{inputenc}
\usepackage[american]{babel}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epsfig}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{color}
\usepackage{datetime}
\usepackage{epstopdf}
\usepackage{hyperref}

\usepackage{csquotes}
\usepackage{listings}
\usepackage[backend=biber]{biblatex}

\usepackage{minted}
\usepackage{float}
\floatstyle{boxed}
\restylefloat{figure}

\addbibresource{refs.bib}

\hypersetup{
    pdfborder = {0 0 0}
}
\newcommand{\orangutan}{Orangutan}
\newcommand{\bibtex}{Bib{\TeX}}
\newcommand{\chapref}[1]{(Section~\ref{#1})}
\newcommand{\rfquote}[2]{
\begin{flushright}
\emph{#1}\\
-- #2
\end{flushright}}

\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\ie}{i.e.}
\newcommand{\eg}{e.g.}
\newcommand{\newdef}[1]{`#1'}
\newcommand{\file}[1]{.#1 file}

\graphicspath{ {graphics/} {./} }
% see http://imf.au.dk/system/latex/bog/

\begin{document}
\pagestyle{empty}
\pagenumbering{Alph}
\begin{titlepage}
\author{Rohde Fischer - 20052356}
\csAdvisor{Olivier Danvy}
\title{An automated approach\\to organizing {\bibtex} files}
\maketitle

\end{titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagestyle{plain}
\setcounter{page}{1}


\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

We successively describe the bibliographic software tool {\bibtex}
(both how to use it in principle and how it is used in practice), list
a range of practical issues {\bibtex} users encounter, propose an
approach to hand\-ling these issues, and review how they are tackled
in related work.  We then present an analysis of {\bibtex} files that
detects these issues and we describe how to solve them by organizing
{\bibtex} files using the results of the analysis.  We implemented a
proof of concept, {\orangutan}, that analyzes existing {\bibtex} files
and emits diagnostics and suggestions.


\chapter*{Resumé}
\addcontentsline{toc}{chapter}{Resumé}

Vi beskriver successivt det bibliografiske software-værktøj {\bibtex}
(både, hvordan det bruges i princippet og hvordan det bruges i
praksis), lister en række praktiske problemer {\bibtex}-brugere møder,
foreslår en fremgangsmåde til håndtering af problemerne og en
gennemgang af, hvordan de takles i relateret arbejde.  Vi præsenterer
en analyse af {\bibtex}-filer, der påviser problemerne og vi
beskriver, hvordan disse kan løses ved at organisere {\bibtex}-filerne
ved hjælp af resultaterne fra analysen.  Vi har implementeret et proof
of concept, {\orangutan}, der analyserer eksisterende {\bibtex}-filer
og udskriver diagnoser og forslag.


\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

Olivier Danvy served as my advisor during the work that led up to this
dissertation.  Deep gratitude goes to his constant support and help,
which have been of immense value.

René Rydhof Hansen served as an external evaluator, and issued
insightful questions at the defense.  I am grateful for his time and attention.

Anders Lindkvist, Troels Fleischer Skov Jensen, and Martin Eik
Korsgaard Rasmussen were fantastic office mates.  They offered
critical feedback on the work, comments on the dissertation, and moral
support throughout.  Thank you guys.

Ditte Maria Mikkelsen, Livia-Marina Gherghe, and Vera Ohlsen
generously provided suggestions and comments on the dissertation.
Thanks for the 



To my advisor, Olivier Danvy, goes a deep thank you.  During the work
that led up to this dissertation, his constant support and help have
been of immense value.

Thanks to René Rydhof Hansen for serving as an external evaluator.

Grateful thanks to my three office mates, Anders Lindkvist, Troels
Fleischer Skov Jensen, and Martin Eik Korsgaard Rasmussen, for
offering critical feedback on the work, giving comments on the
dissertation, and providing moral support throughout.

Thanks to Ditte Maria Mikkelsen, Livia-Marina Gherghe, and Vera Ohlsen
for generously providing suggestions and comments on the dissertation.

I also want to thank the JabRef community its responsiveness.

My gratitude also goes to the staff at Aarhus University for providing
a fruitful study environment.  Through their diligent work, they have
been ensuring the relevant infrastructure for my studies, from
maintaining the computer systems to assisting with practical issues.

To Bookstudenterkørsel.dk a huge thanks is due for providing a student
job having an understanding and being accepting of the requirements for
my studies.  My boss and colleagues have been providing a perfect
unification of both studies and work.

I am also grateful to my siblings, Karl and Anne, and to my sister in
law, Julija, for the moral support and all the help and feedback they
tirelessly provided.

To my good friends Peter Laursen, Iskra Dinkova, Livia-Marina
Gherghe, and Nina Sprogøe another thanks for the moral support they
graciously provided.

To Kim Andersen, Rasmus Langager Berg, and Michael Lycke thanks are
also due for helping me with my practicalities.


\tableofcontents
\newpage

\pagenumbering{arabic}
\setcounter{secnumdepth}{2}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}
\label{ch:intro}
\input{intro.tex}

\chapter{{\bibtex}}
\label{ch:about}
\input{about.tex}

\chapter{The challenges\\in using {\bibtex}}
\label{ch:problem-description}
\input{problem-description.tex}

\chapter{Our approach to {\bibtex}}
\label{ch:approach}
\input{approach.tex}

\chapter{Related work}
\label{ch:related}
\input{related.tex}

\chapter{Analyzing {\bibtex} files}
\label{ch:analyzing}
\input{analyzing.tex}

\chapter{Organizing {\bibtex} files}
\label{ch:organizing}
\input{organizing.tex}

\chapter{Conclusion and perspectives}
\label{ch:conclusion}
\input{conclusion.tex}

\printbibliography{}
\end{document}
